---
title: "  **ALY6030 : Analysis of Cryptocurrencies**"
output: pdf_document
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{setspace}\spacing{1.5}
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=6in,height=11in]{image.png}\LARGE\\}
  - \posttitle{\end{center}}
  - \usepackage{sectsty} \sectionfont{\emph}
---
\begin{centering}
\vspace{1cm}
\Large
{\bf Professor : Dr.Hema Seshadri} \\
\vspace{.75cm}
\large
{\bf Submitted By : Sharon Appoline Rosary and David Joseph Johnson} \\
\vspace{.75cm}
{\bf 05/17/2022 } \\
\end{centering}
\newpage
\tableofcontents
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This file contains the R code used to Clean and create visualizations of the dataset.


## Loading Libraries

```{r , echo=FALSE ,results='hide', warning=FALSE, message=FALSE}
### Install ISLR Package which contains the College Dataset
### Invoke Packages
library("ISLR")
library("ggplot2")
library("corrplot")
library("dplyr")
library("janitor")
library("tidyr")
library("RColorBrewer")
library("psych")
library("imputeTS")
library("corrplot")
library("leaps")
library("knitr")
library("tibble")
library("stargazer")
library("lattice")
library("caret")
library("pROC")
library("glmnet")
library("gridExtra")
library("stringr")
```

## Data Cleaning
```{r , echo=FALSE ,results='hide', warning=FALSE, message=FALSE}
raw_data <- read.csv(
  "data.csv",
  header = TRUE)

# Count of records before cleaning
nrow(raw_data)

# Columns in Raw Data
names(raw_data)

#Drop Unwanted columns
raw_data <- raw_data[-c(4,11,13:16,29)]

# Columns in Raw Data after Drop
names(raw_data)

#Remove empty rows and columns.
raw_data <- remove_empty(raw_data,which = c("rows","cols"),quiet=TRUE)

#Check for duplicate records.
get_dupes(raw_data)

raw_data['mineable'] <- c("")
raw_data['exchange'] <- c("")
raw_data['payments'] <- c("")
names(raw_data)

# Update Tags
for (i in 1:nrow(raw_data))
  {
  ## Initializing -100000 to records with unknown or infinite "max supply"
  raw_data[i,7] = if (is.na(raw_data[i,7]) == TRUE) -100000 else raw_data[i,7]  
  
  ## Initializing values to mineable exchange and payments column
  raw_data[i,23] = if ((str_detect(raw_data[i,6], "mineable")) == TRUE) 'Yes' else 'No'
  raw_data[i,24] = if ((str_detect(raw_data[i,6], "exchange")) == TRUE) 'Yes' else 'No'
  raw_data[i,25] = if ((str_detect(raw_data[i,6], "payments")) == TRUE) 'Yes' else 'No'
}

# Remove Tags
raw_data <- raw_data[-c(6)]

#Check for NAs
knitr::kable(sum(is.na(raw_data)),col.names = "No. of NAs")

write.csv(raw_data,"cleanCryptoData.csv",row.names = FALSE)
```

# Analysis

## Overview of Dataset

Initial Analysis is performed to better understand the dataset and variables under observation,the HEAD and TAIL of the dataset, and the structure of dataset is looked-up.

```{r }
df<-raw_data
# Head,Tail and Structure of Dataset
head(df)
tail(df)
str(df)
describe(df)

top_ranked <- head(df)
bottom_ranked <- tail(df)
```

## Barplot of Year in which Cryptocurrency was added
```{r }

df$year <- substr(df$date_added, 0, 4)

ggplot(data=df, aes(x=year)) +
  geom_bar(stat="count", fill="steelblue")+
  geom_text(stat='count', aes(label=..count..), vjust=-0.5)+
  xlab("Year Added") + ylab("Count of Cryptocurrencies")+
  theme_minimal()


```


## Top 10 Most Expensive Crypto

```{r }



## OrderBy Price (DESC)
df2 <- df[order(-df$price),]

## Filter Top 10
priceDesc <- top_n(df2,10,df2$price)

p <- ggplot(priceDesc, aes(x = reorder(name, -price), y = price)) + geom_bar(stat="identity", color='skyblue',fill='steelblue') +
  labs(y="Price", x="Cryptocurrency", title="Top 10 Most Expensive Crypto")+
theme(axis.text.x=element_text(angle=90, hjust=1))+ 
  geom_text(label=round(priceDesc$price,digit=1),position=position_dodge(width=0.3),
                                                 hjust=1.1,angle=90,vjust=1)
p
```


## Count of Tags

```{r }
catCount <-c()
for (i in 1:nrow(df))
  {
  if(identical(df[i,22],"Yes")){
    catCount <- append(catCount,"MINEABLE")
  }
 if(identical(df[i,23],"Yes")){
    catCount <- append(catCount,"EXCHANGE")
  }
  if(identical(df[i,22],"Yes")){
    catCount <- append(catCount,"PAYMENTS")
  }
  
}

q <- ggplot(data.frame(catCount), aes(x=catCount)) +
  geom_bar(stat="count", fill="steelblue")+
  geom_text(stat='count', aes(label=..count..), vjust=-0.5)+
  labs(y="COUNT", x="TAG", title="Categories")
q
```


## Scatterplot of Price Vs Total Supply of currencies in Lower Price Range

```{r }

ggplot(data=df,aes(total_supply,price)) +
   geom_point(color="slateblue1") + 
   ggtitle("Scatterplot of Price Vs Total Supply of currencies in Lower Price Range") + labs(y="Price",x= "total_supply")+ ylim(c(0,1000)) + xlim(c(0,500000))

```

## Jitterplot of Total Supply Vs Mineable

```{r }


ggplot(df, aes(x=mineable, y=total_supply/1000000, color=mineable)) + 
  geom_boxplot(notch=FALSE,outlier.colour="yellow", outlier.shape=8,
                outlier.size=1) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  labs(caption="Jitterplot of Total Supply Vs Mineable",
  x="Is Mineable ?", y = "Supply (in Billions)?") +
  theme(plot.caption = element_text(hjust = 0.5))+ylim(c(0,100000))


```

## ANOVA Test

```{r }

# One-way ANOVA Test

# Set significance level
alpha <- 0.05
# Dataframe for 30 Days
thirtyDays <- data.frame('variation' = df$percent_change_30d,
                         'timeperiod' = rep('thirtyDays',5000), stringsAsFactors = FALSE)
# Dataframe for 60 Days
sixtyDays <- data.frame('variation' = df$percent_change_60d,
                      'timeperiod' = rep('sixtyDays',5000), stringsAsFactors = FALSE)
# Dataframe for 90 Days
ninetyDays <- data.frame('variation' = df$percent_change_90d,
                       'timeperiod' = rep('ninetyDays',5000), stringsAsFactors = FALSE)
# Combine the Dataframe
variation <- rbind(thirtyDays,sixtyDays,ninetyDays)
variation$timeperiod <- as.factor(variation$timeperiod)

# Hypotheses
# H0: Mean Variation(30 Days) =  Mean Variation(60 Days) =  Mean Variation(90 Days)
# H1: Atleast one mean is different from others

anova <- aov(variation ~ timeperiod, data = variation)
a_summ <-summary(anova)
# Critical Value
qf(1-alpha,a_summ[[1]][1,1],a_summ[[1]][2,1])

# Test Value
F.value <- a_summ[[1]][[1,"F value"]]
F.value

# Compare p-value and alpha to make decision
p.value <- a_summ[[1]][[1,"Pr(>F)"]]
p.value
ifelse(p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
```

## Correlation Analysis of Numeric values

```{r }
numeric_data <- select_if(df, is.numeric) 
data.cor = cor(numeric_data, method = c("spearman"))
corrplot(data.cor)

```


